### 2023-02-01 manuscript.Rmd
shorten
typo

# Aspects of rigor for Software Implementations
Best-practises for software development can be used to assess rigor for research software.
In the following, we discuss aspects that should be scored in phase one of the assessment.
For the reasons outline above, we propose to score those instead of the reusability indicator (item 17 in table 3).
<!-- some introductory text missing -->
<!-- maybe cite
https://joss.readthedocs.io/en/latest/review_criteria.html
and
https://joss.readthedocs.io/en/latest/review_checklist.html
here -->
## Tests
Tests are essential to verify the correctness of scientific software.
They are usually separated into different categories:
*Unit tests* evaluate only small pieces of code, for example specific R-functions.
*Functional Tests* test whether the input of a user interacting with the software leads to the correct output.
There are many more types of tests, and the type of tests appropriate for a specific software product depends on the software itself.
However, we argue it should at least be tested whether the software can be installed on different platforms, and whether the core functionality of the software is executed without errors (*Smoke Test*).
In addition, it might be possible to give higher scores if different types of tests are employed, or to define a minimum of tests types that should be present, for example *unit tests*, *integration tests* and *functional tests*.
Another metric that could be assessed is code coverage, defined as the percentage of code lines that are executed during testing.
However, we would like to note that this metric is to be taken with a grain of salt, as the quality and number of the tests is not taken into account.
Tests should be automated, or at least reproducible.
This rules out closed-source software tests.

@@

## Peer-Review

Similar to peer-review for manuscripts, software code is amenable to peer-review.
It is common practice in software development to have peers review code.
For example, if a code base is changed, it is good practice to request a change with the developer team before actually changing the code base.
The change is reviewed by experts and/or maintainers before it is integrated in the active branch of development.
In developing these guidelines, we see the opportunity to foster such good practices and reward researchers that actively engage in reviewing computer code.
Software peer-review is increasingly becoming recognized as a means to enable better research software.
Peer-review can happen in the context of dedicated software journal (e.g, Journal of Statistical Software), as part of the standard review process (for guidelines, see https://www.carlboettiger.info/2013/06/13/what-I-look-for-in-software-papers.html), or as part of submissions to software archives that guarantee certain standards (e.g., rOpenSci).
We suggest that active participation in the review of scientific software, independent of the involvement or maintenance of specific software products, should be a dimension of positive evaluation.

>>>>

# Aspects of rigor for Software Implementations
Best-practises for software development can be used to assess rigor for research software.
In the following, we discuss aspects that should be scored in phase one of the assessment.
For the reasons outline above, we propose to score those instead of the reusability indicator (item 17 in table 3).
<!-- some introductory text missing -->
<!-- maybe cite
https://joss.readthedocs.io/en/latest/review_criteria.html
and
https://joss.readthedocs.io/en/latest/review_checklist.html
here -->

## Tests
Tests are essential to verify the correctness of scientific software.
They are usually separated into different categories:
*Unit tests* evaluate only small pieces of code, for example specific R-functions.
*Functional Tests* test whether the input of a user interacting with the software leads to the correct output.
There are many more types of tests, and the type of tests appropriate for a specific software product depends on the software itself.
However, we argue it should at least be tested whether the software can be installed on different platforms, and whether the core functionality of the software is executed without errors (*Smoke Test*).
In addition, it might be possible to give higher scores if different types of tests are employed, or to define a minimum of tests types that should be present, for example *unit tests*, *integration tests* and *functional tests*.
Another metric that could be assessed is code coverage, defined as the percentage of code lines that are executed during testing.
However, we would like to note that this metric is to be taken with a grain of salt, as the quality and number of the tests is not taken into account.
Tests should be automated, or at least reproducible.
This rules out closed-source software tests.

@@

## Peer-Review
Similar to peer-review for manuscripts, software code is amenable to peer-review.
It is common practice in software development to have peers review code.
For example, if a code base is changed, it is good practice to request a change with the developer team before actually changing the code base.
The change is reviewed by experts and/or maintainers before it is integrated in the active branch of development.
In developing these guidelines, we see the opportunity to foster such good practices and reward researchers that actively engage in reviewing computer code.
Software peer-review is increasingly becoming recognized as a means to enable better research software.
Peer-review can happen in the context of dedicated software journal (e.g, Journal of Statistical Software), as part of the standard review process (for guidelines, see https://www.carlboettiger.info/2013/06/13/what-I-look-for-in-software-papers.html), or as part of submissions to software archives that guarantee certain standards (e.g., rOpenSci).
We suggest that active participation in the review of scientific software, independent of the involvement or maintenance of specific software products, should be a dimension of positive evaluation.
